---
title: "Untitled"
output:
  word_document: default
  html_document: default
date: "2023-01-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load standard libraries

```{r}
library(tidyverse) #data wrangling and data visualization
library(caTools)#splitting data into training and testing sets
library(pROC)#measuring and comparing diagnostic test performance
library(randomForest)#building predictive models
library("readr") #importing and exporting flat files
library("ggplot2")#data visualization 
library("dplyr")#data manipulation
library("caret")#machine learning algorithms
library(corrgram)#visualizing correlations

```
# Setting the working directory and Loading the data set
```{r}
setwd("D:/CSDA")

bloodTransfusion<- read.csv("Blood Transfusion Dataset (1).csv")

head(bloodTransfusion)


```


#Changing Column name for target variable
```{r}
colnames(bloodTransfusion)[colnames(bloodTransfusion)=="whether.he.she.donated.blood.in.March.2007"]<-"DonatedMar07"
#Transforming target variable to factors
bloodTransfusion$DonatedMar07 <- as.factor(bloodTransfusion$DonatedMar07)

str(bloodTransfusion)
#Renaming the column names for easier use/reference

n <- c("Recency", "Frequency", "Monetary", "Time", "DonatedMar07")
colnames(bloodTransfusion) <- n

```



##### Summary statistics
```{r}

summary(bloodTransfusion)

```


##Exploratory data analysis
###Boxplot of the Months since Last Donation
```{r}
##BOXPLOTS
ggplot(bloodTransfusion, aes(x = "", y = Time)) +
  geom_boxplot(color = "blue", fill = "turquoise") +
  ylab("Months since Last Donation") +
  ggtitle("Boxplot of the Months since Last Donation") 

```


##Boxplot of the Months since Last Donation across the response



```{r}
ggplot(bloodTransfusion, aes(x=DonatedMar07, y=Time,color=DonatedMar07)) + 
  geom_boxplot(aes(fill=DonatedMar07)+ facet_grid(. ~ DonatedMar07))+
  xlab("DonatedMar07") + ylab("Boxplot of the Months since Last Donation across the response") +theme_bw()

```



####Scatter plots
```{r}
library(ggplot2)
ggplot(data = bloodTransfusion, aes(x = Time, y =Recency)) +
  geom_point() +
  geom_smooth(method=lm)

```



####Scatter plots
```{r}
library(ggplot2)
ggplot(data = bloodTransfusion, aes(x = Time, y =Frequency)) +
  geom_point() +
  geom_smooth(method=lm)
```


####Density plot  to show distribution

```{r}
ggplot(bloodTransfusion, aes(x=Time)) + 
  geom_density() + 
  theme(legend.position="none") + 
  xlab("Months since Last Donation") + 
  ggtitle("Months since Last Donation Density") + 
  geom_vline(aes(xintercept=mean(Time)), color="blue", linetype="dashed", size=1)
```



```{r}

ggplot(bloodTransfusion, aes(x=Monetary)) + 
  geom_density() + 
  theme(legend.position="none") + 
  xlab("Total Volume of Blood Donated") + 
  ggtitle("Total amount  of blood that the donor has donated in C.C") + 
  geom_vline(aes(xintercept=mean(Monetary)), color="blue", linetype="dashed", size=1)


```

####Summary statistics
```{r}

summary(bloodTransfusion$Time)

```




#####Barplots showing frequencies
```{r}

ggplot(bloodTransfusion, aes(DonatedMar07)) +
  geom_bar(stat = "count", aes(fill = DonatedMar07)) + 
  ggtitle("Distribution of Response variable") + 
  theme(legend.position="none")

```


###Correlation coefficient showing association
panel.cor <- function(x, y, digits = 2, cex.cor, ...)
{
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(0, 1, 0, 1))
  # correlation coefficient
  r <- cor(x, y)
  txt <- format(c(r, 0.123456789), digits = digits)[1]
  txt <- paste("r= ", txt, sep = "")
  text(0.5, 0.6, txt)
```{r}

corrgram(bloodTransfusion, order=NULL, lower.panel=panel.shade, upper.panel=NULL, text.panel=panel.txt,
         main="Corrgram of the data")

```
 

##MODELLING

###Training and testing data sets
### Split the data into a training and testing set


```{r}

# Set seed for reproducibility
set.seed(102928)

#splits the data in the ratio mentioned in SplitRatio. After splitting marks these rows as logical 
# TRUE and the the remaining are marked as logical FALSE
sample <- sample.split(bloodTransfusion, SplitRatio = .75)

# creates a training dataset named train with rows which are marked as TRUE
training <- subset(bloodTransfusion, sample == TRUE)

# creates a training dataset named test with rows which are marked as FALSE
testing  <- subset(bloodTransfusion, sample == FALSE)

dim(training) 
dim(testing)


```

##Random Forest Algorithm

```{r}
set.seed(2304) 
modFit_RF = train(DonatedMar07 ~ ., method = "rf", data = training, prox = TRUE, na.action=na.omit)
modFit_RF
plot(modFit_RF)
```



####Variable Importance

```{r}
rf_imp = varImp(modFit_RF, scale = FALSE)
rf_imp
plot(rf_imp)
```

###Model validation

```{r}
pred_rf = predict(modFit_RF, testing)
confusionMatrix(pred_rf, testing$DonatedMar07)

```



####Applying model to the test sample.
```{r}
test_sample = bloodTransfusion

test_sample$pred_rf = predict(modFit_RF, test_sample, type = "prob")[,2]


ggplot(test_sample, aes(x=pred_rf)) + 
  geom_density() + 
  theme(legend.position="none") + 
  xlab("Random Forest Probability") + 
  ggtitle("Probability that they donated in March") + 
  geom_vline(aes(xintercept=mean(pred_rf)), color="blue", linetype="dashed", size=1)


```


####Summary of probability 
```{r}

summary(test_sample$pred_rf)
```





##Logistic Regression

```{r}

set.seed(2304) 
modFit_lr = train(DonatedMar07 ~ ., data = training, method = "glm", family = "binomial", na.action=na.omit)
modFit_lr

lr_imp = varImp(modFit_lr, scale = FALSE)
lr_imp

plot(lr_imp)

```


##Generating predictions from Logistic model

```{r}
pred_lr = predict(modFit_lr, testing)
confusionMatrix(pred_lr, testing$DonatedMar07)



test_sample$pred_lr = predict(modFit_lr, test_sample, type = "prob")[,2]

ggplot(test_sample, aes(x=pred_lr)) + 
  geom_density() + 
  theme(legend.position="none") + 
  xlab("Logistic Regression Probability") + 
  ggtitle("Probability that they donated in March  using Logistic regression") + 
  geom_vline(aes(xintercept=mean(pred_lr)), color="blue", linetype="dashed", size=1)


summary(test_sample$pred_lr)

```

